---
title: "Untitled"
output: html_document
---

```{r}
library(tsibble)
library(tsibbledata)
library(fable)
library(fabletools)
```

```{r}
tsibbledata::vic_elec
```



## Mean model

$$
    \hat{y}_{T+h|T} = \bar{y} = (y_1 + y_2 + \dots + y_T)/T
$$

```{r}
mean_mod <- tsibbledata::vic_elec %>%
  fabletools::model(mean = fable::MEAN(Demand))
```

```{r}
mean_mod
```

```{r}
mean_mod %>% fabletools::forecast(h=90)
```

```{r}
mean_mod %>% fabletools::forecast(h="90 days")
```

```{r}
mean_mod %>% fabletools::forecast(h="90 days") %>%
  fabletools::autoplot(
    #dplyr::filter(tsibbledata::vic_elec, Time >= as.Date("2014-01-01"))
    tsibbledata::vic_elec
  )
```

## Naive

Just the last value

$$
    \hat{y}_{T+h|T} = y_T
$$

```{r}
naive_mod <- tsibbledata::vic_elec %>% 
    model(Naive=NAIVE(Demand))

naive_mod

naive_mod %>% forecast(h=90)
naive_mod %>% forecast(h="90 days")

naive_mod %>% forecast(h="90 days") %>% autoplot(tsibbledata::vic_elec)
```

## Seasonal Naive

The forecast is the last observed value from the same seasons of the year.

$$
\hat{y}_{T+h|t} = y_{T+h-m(k+1)}
$$

where $m$ is the seasonal period and $k$ is the number of complete years in the forecast period.

```{r seasonal-naive-mod,dependson=c('make-tsibble')}
snaive_mod <- tsibbledata::vic_elec %>% 
    model(SNaive=SNAIVE(Demand ~ lag("year")))

snaive_mod

snaive_mod %>% forecast(h=90)
snaive_mod %>% forecast(h="90 days")

snaive_mod %>% forecast(h="90 days") %>% autoplot(tsibbledata::vic_elec)
```

Reflects seasonal changes.

## Drift Method

Last value plus average change over all the data

Random Walk with Drift

$$
    \hat{y}_{T+h|T} = y_T + \frac{h}{T-1}\sum_{t=2}^T (y_t - y_1) = y_T + h \left( \frac{y_T - y_1}{T - 1} \right)
$$

```{r drift-mod,dependson=c('make-tsibble')}
drift_mod <- tsibbledata::vic_elec %>% 
    model(Drift=RW(Demand ~ drift()))

drift_mod

drift_mod %>% forecast(h=90)
drift_mod %>% forecast(h="90 days")

drift_mod %>% forecast(h="90 days") %>% autoplot(tsibbledata::vic_elec)
```

## Compare the Methods

See all of them on one plot

```{r}
simple_mods <- tsibbledata::vic_elec %>% 
    model(
        Mean=MEAN(Demand)
        , Naive=NAIVE(Demand)
        , SNaive=SNAIVE(Demand ~ lag("1 year"))
        , Drift=RW(Demand ~ drift())
    )

simple_mods

simple_mods %>% forecast(h="90 days")
simple_mods %>% 
    forecast(h="90 days") %>% 
    autoplot(filter_index(tsibbledata::vic_elec, '2010-09' ~ .), level=NULL)
```

## prediction intervals

```{r}
simple_mods %>%
  fabletools::forecast(h="90 days") %>%
  fabletools::hilo(level = c(50, 75)) %>%
  tidyr::unnest()
```

