---
title: "Untitled"
output: html_document
---

```{r}
library(fable)
library(tsibble)
library(tsibbledata)
```


Exponentially weighted moving average

$$
\hat{y}_{T+1|T} = \alpha y_T + \alpha(1 - \alpha)y_{T-1} + \alpha(1 - \alpha)^2y_{T-2} + \cdots
$$

If $\ell_0$ is the first observation then this generalizes to

$$
\hat{y}_{T+1|T} = \sum_{j=0}^{T-1} \alpha(1 - \alpha)^j y_{T-j} + (1 - \alpha)^T \ell_0
$$

In component form this is

$$
\begin{align}
\text{Forecast Equation: } & \hat{y}_{t+h|t} = \ell_t \\
\text{Smoothing Equation: } & \ell_t = \alpha y_t + (1 - \alpha) \ell_{t-1}
\end{align}
$$

linear trend, AAN MAN

```{r}
us_eco <- global_economy %>%
  dplyr::filter(Code == "USA") %>%
  dplyr::mutate(pop = Population / 1e6)
```

```{r}
ets_aan <- us_eco %>%
  fabletools::model(
    aan = fable::ETS(pop) # can provide specials
  )
```

```{r}
ets_aan
```

```{r}
fabletools::report(ets_aan) # not for inference
```


```{r}
components(ets_aan)
```

```{r}
components(ets_aan) %>% autoplot()
```

```{r}
ets_aan %>%
  forecast(h = 20) %>%
  autoplot(us_eco)
```

## aan

```{r}
aan <- us_eco %>%
  fabletools::model(
    ann = fable::ETS(pop ~ error("A") + trend("A") + season("N"))
  )

aan %>%
  forecast(h = 30) %>%
  autoplot(us_eco)
```

## aaa

```{r}
aaa <- us_eco %>%
  fabletools::model(
    ann = fable::ETS(pop ~ error("A") + trend("A") + season("A"))
  )

aaa %>%
  forecast(h = 30) %>%
  autoplot(us_eco)
```

## Automatic model formulation

Best determined by $AICc$

```{r ets-auto-model,dependson=c('make-tsibble')}
ets_auto <- us_eco %>% 
  model(
    ets=ETS(pop)
  )

ets_ann %>% report

ets_auto %>% 
  forecast(h=30) %>% 
  autoplot()
```

# AICc

$$
\text{AIC} = -2 \text{log}(L) + 2(p+q+k+1)
$$

$$
\text{AICc} = \text{AIC} + \frac{2(p+q+k+1)(p+q+k+2)}{T - p - q - k - 2}
$$

## Exercise

1. Try forcasting the USA GDP
2. Do it with with a boxcox transformation

```{r}
lambda <- us_eco %>% 
   features(GDP, features=feasts::guerrero) %>% 
   dplyr::pull(lambda_guerrero)

lambda

us_eco %>% 
    autoplot(box_cox(GDP, lambda))

ets <- us_eco %>% 
  model(
    ets=ETS(GDP)
  )

ets %>% report

ets %>% 
  forecast(h=30) %>% 
  autoplot()

