---
title: "Untitled"
output: html_document
---


# Autoregressive Models


$$
y_t = c+ \phi_1y_{t-1} + \phi_2y_{t-2} + \cdots + \phi_py_{t-p} + \epsilon_t
$$

# Moving Average Models

$$
y_t = c + \epsilon_t + \theta_1\epsilon_{t-1} + \theta_2\epsilon_{t-2} + \cdots + \theta_q\epsilon_{t-q}
$$

# ARIMA Models

Autoregressive Integrated Moving Average

$$
y_t' = c + \phi_1y_{t-1}' + \cdots + \phi_py_{t-p}' + \theta_1\epsilon_{t-1} + \cdots + \theta_q\epsilon_{t-q} + \epsilon_t
$$

AR: p = order of the autoregressive part
I: d = degree of rst dierencing involved
MA: q = order of the moving average part.

White noise model: ARIMA(0,0,0)
Random walk: ARIMA(0,1,0) with no constant
Random walk with drift: ARIMA(0,1,0) with const.
AR(p): ARIMA(p,0,0)
MA(q): ARIMA(0,0,q)

```{r}
library(tsibble)
library(tsibbledata)
library(fable)
library(dplyr)

elec <- readr::read_csv(here::here("data", "electricity_france.csv"))
elec <- elec %>% 
    as_tsibble(index=Date) %>% 
    filter(Date >= as.Date('2007-01-01'))
elec
```

Auto select a model

pdq special is used to specify non-seasonal components of the model

PDQ special is used to specify seasonal components of the model

```{r}
elec_arima_000 <- elec %>% 
  model(
    ARIMA(ActivePower ~ PDQ(0, 0, 0))
  )
```

```{r}
elec_arima_000
elec_arima_000 %>% report()

```

```{r}
elec_arima_000 %>% 
  forecast(h=60) %>% 
  autoplot() + 
  autolayer(filter_index(elec, '2010-10' ~ .), ActivePower)
```

acf and pacf plots

```{r}
elec %>% feasts::ACF(ActivePower) %>% autoplot()
elec %>% feasts::PACF(ActivePower) %>% autoplot()
```

arima auto harder

```{r}
elec_arima_auto <- elec %>% 
  model(
    ARIMA(ActivePower ~ PDQ(0, 0, 0), 
          stepwise = FALSE, approximation = FALSE
    )
  )

elec_arima_auto
elec_arima_auto %>% report()

elec_arima_auto %>% 
  forecast(h=60) %>% 
  autoplot(filter_index(elec, '2010-10' ~ .))
```

```{r}
elec_arima_auto_2 <- elec %>% 
  model(
    ARIMA(ActivePower ~ pdq(1:4, 1:2, 1:4) + PDQ(0, 0, 0), 
          stepwise = FALSE, approximation = FALSE
    )
  )

elec_arima_auto_2
elec_arima_auto_2 %>% report()

elec_arima_auto_2 %>% 
  forecast(h=60) %>% 
  autoplot(filter_index(elec, '2010-10' ~ .))
```

auto arima

```{r}
elec_auto_all <- elec %>% 
  model(
    arima=ARIMA(ActivePower)
  )

elec_auto_all %>% report()
# Produces a plot of the inverse AR and MA roots of an ARIMA model. Inverse roots outside the unit circle are shown in red.
# Only models which compute ARMA roots can be visualised with this function
elec_auto_all %>% feasts::gg_arma()
```

```{r}
elec_auto_all %>% forecast(h=60)
elec_auto_all %>% 
  forecast(h=30) %>% 
  autoplot(filter_index(elec, '2010-06' ~ .))
```

